---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r}
library(MALDIquantForeign)
library(MALDIquant)
library(tidyverse)
library(parallel)
library(cowplot)
library(ggpubr)
library(ggrepel)
library(sva)
library(binda)
library(sda)
library(limma)
library(pheatmap)
library(RColorBrewer)
library(viridis)

backup_options = options()
backup_par = par()
```


# Data analysis of the preprocessed MALDI data

We analyse the preprocessed MALDI data in the form of feature matrices, where samples are the rows and each feature/mass/peak/peptide the columns; the matrix cells are filled with the processed intensity values.

```{r}
basepath = '~/palaeoproteomics/MALDI/spanish_book'
data_folder = file.path(basepath, 'data')
results_folder = file.path(basepath, 'results')

# -----------------------------------------------------------------------------------------
# Read metadata
uoc_metadata = read.csv(file.path(data_folder,'uoc_metadata.csv'))
uoc_metadata = uoc_metadata %>%
  mutate(group = as.factor(group), book = as.factor(book), side = as.factor(side),
         batch = as.factor(batch)) %>% as_tibble()

# -----------------------------------------------------------------------------------------
# Read feature matrix
snr = 5
featMatrix = read_csv(file.path(data_folder,
                                paste0('featureMatrix_',snr,'.csv')),
                      show_col_types = F) %>%
  as.matrix()
colnames(featMatrix) = as.character(round(as.numeric(colnames(featMatrix)),3))
rownames(featMatrix) = uoc_metadata$sample.name
# Scale
scaledFeatMatrix = apply(featMatrix, 2, function(x) (x)/sqrt(var(x)))
```


## Surrogate varaible analysis
First of all we will use the `sva` package to estimate and regress out the surrogate variables. For this, we will tell the algorithm which are our variables of interest, _species_ and _book_ and we will let it determine unwanted sources of variation, among which we have the batch.


```{r}
# Prepare models for SVA analyis
mod = model.matrix(~species+group, data = uoc_metadata)
mod0 = model.matrix(~1, data=uoc_metadata)
# Remove surrogate variables
# sva
n.sv = num.sv(t(scaledFeatMatrix), mod, method='be', B=47)
svobj = sva(t(scaledFeatMatrix), mod, mod0, n.sv=n.sv)
colnames(svobj$sv) = paste0('SV', 1:n.sv)
# Fix data
brSVA = fsva(dbdat=t(scaledFeatMatrix), mod=mod, sv=svobj, newdat=t(scaledFeatMatrix))
brSVA = brSVA$db
```

We can calculate F-test for each of the masses to have a preliminary idea of whether they have differences in the different species and books
```{r, fig.width = 9, fig.height = 6}
# Add surrogate variables to model and estimate pvalues
modSV = cbind(mod, svobj$sv)
mod0SV = cbind(mod0, svobj$sv)
# Calc fstatistics
pValuesSv = f.pvalue(t(scaledFeatMatrix), modSV, mod0SV)
qValuesSv = sort(p.adjust(pValuesSv, method="BH"))

sv_qval = tibble(mass = as.numeric(names(qValuesSv)), q_val = qValuesSv)
sv_qval %>% filter(q_val < 0.05)
pl = ggplot(sv_qval) +
    geom_point(aes(x=mass, y=-log10(q_val)), size=2, alpha=0.7) +
    geom_hline(yintercept=-log10(0.05), color='red') +
    geom_text_repel(aes(x=mass, y=-log10(q_val), label=ifelse(q_val<0.05, mass, '')), size=4) +
    theme_cowplot() +
    theme(axis.title=element_text(size=12), axis.text=element_text(size=10))
pl
ggsave('f_stats_pval.png', pl, path = results_folder, bg='white',
       width = 9, height = 6, dpi=100)

ggsave('f_stats_pval.svg', pl, path = results_folder, bg='white',
       width = 9, height = 6, dpi=100)

```


## Data visualization and clustering
We perform a hierarchical clustering on the samples and a PCA.

### Hierarchical cluster
```{r, fig.width = 12, fig.height = 6}
# Hierarchical cluster
h = hclust(dist(t(brSVA)), method='ward.D2')
png(file.path(results_folder,'hier_clust.png'),
    width = 12, height = 6, units = 'in', res = 100)

svg(file.path(results_folder,'hier_clust.svg'),
    width = 12, height = 6)
a = plot(h, labels=paste0(uoc_metadata$folia, uoc_metadata$side, '-', uoc_metadata$group, '-',
                      uoc_metadata$species, collapse = NULL),
     ylab='Distance', xlab='Sample', axes=FALSE, sub='', main='Hierarchical clustering',
     cex=1
    )
dev.off()

plot(h, labels=paste0(uoc_metadata$folia, uoc_metadata$side, '-', uoc_metadata$group, '-',
                      uoc_metadata$species, collapse = NULL),
     ylab='Distance', xlab='Sample', axes=FALSE, sub='', main='Hierarchical clustering',
     cex=1
    )
```

### Heatmap

```{r, fig.width = 9, fig.height = 7}

annot = uoc_metadata %>%
    column_to_rownames('sample.name') %>%
    as.data.frame() %>%
    select(species, book)
cl = paste0(uoc_metadata$folia, uoc_metadata$side, '-', uoc_metadata$group, collapse = NULL)

ann_colors = list(
    species = c(Goat = "#CC6D68", Sheep = "#DBBD68"), 
    book = c("1" = "#238A8DFF", "2" = "#4B2991")
)
col2 = colorRampPalette(c("white","#EACDC2", "#B75D69", "#774C80", "#372549", "#1A1423"))

hm = pheatmap(
  brSVA, annotation_col = annot, annotation_colors = ann_colors,
  color=col2(8), treeheight_row = 0, show_rownames = F,
  clustering_method = 'ward.D2', labels_col = cl, cutree_col=5,
  annotation_names_col = F, fontsize = 8, cex=1)
hm
ggsave('heatmap.png', hm, path=results_folder, width = 9, height = 7,
       dpi=100, bg='white')
ggsave('heatmap.svg', hm, path=results_folder, width = 9, height = 7,
       dpi=100, bg='white')

#flip plot
hm_fl = pheatmap(
  t(brSVA), annotation_row = annot, annotation_colors = ann_colors,
  color=col2(8), treeheight_col = 0, show_colnames = F,
  clustering_method = 'ward.D2', labels_row = cl, cutree_row=5,
  annotation_names_row= F, fontsize = 8, cex=1)
hm_fl
ggsave('heatmap_flipped.png', hm_fl, path=results_folder, width = 9, height = 7,
       dpi=100, bg='white')
ggsave('heatmap_flipped.svg', hm_fl, path=results_folder, width = 9, height = 7,
       dpi=100, bg='white')
```

### PCA
```{r, fig.width = 7, fig.height = 5}
masses = as.numeric(colnames(scaledFeatMatrix))
peptides = c(2883, 3017, 3033, 3077, 3093)

mask1 = masses > peptides[1] & masses < peptides[1]+1
mask2 = masses > peptides[2] & masses < peptides[2]+1
mask3 = masses > peptides[3] & masses < peptides[3]+1
mask4 = masses > peptides[4] & masses < peptides[4]+1
mask5 = masses > peptides[5] & masses < peptides[5]+1


batch_rem_pc = prcomp(t(brSVA))
batch_rem_pc_x = as.data.frame(batch_rem_pc$x)
batch_rem_pc_x = cbind(batch_rem_pc_x, uoc_metadata)
perc_var = round(batch_rem_pc$sdev^2*100/sum(batch_rem_pc$sdev^2), 2)

loadings = batch_rem_pc$rotation[, c(1,2,3)] * batch_rem_pc$sdev[c(1,2,3)]
toppc1_loadings = loadings[order(loadings[,'PC1'], decreasing=T),]
plot_loadings = loadings[mask1| mask2 | mask3 | mask4 | mask5, ] * 10

mass_labels = masses[mask1 | mask2 | mask3 | mask4 | mask5]

# PC1 vs PC2
p1 = ggplot(batch_rem_pc_x) +
  geom_point(mapping = aes (x=PC1, y=PC2, color=species),
             size=6, shape=16, stroke=2) +
  xlab(paste0('PC1. ', perc_var[1], "% variance explained")) +
  ylab(paste0('PC2. ', perc_var[2], "% variance explained")) +
  scale_color_manual(values= c(Goat = "#CC6D68", Sheep = "#DBBD68")) + 
  geom_text(mapping = aes(x=PC1, y=PC2, label=as.character(group)),
            color="white", size=3) +
  theme_cowplot()
p1

p2 = p1 +
  geom_segment(lineend = "butt", linejoin = "round", arrow = arrow(length = unit(0.1, "inches")),
              mapping=aes(xend=PC1,yend=PC2,x=rep(0,5), y=rep(0,5)),
              data=as.data.frame(plot_loadings)) +
  geom_text_repel(mapping=aes(x=PC1, y=PC2,label=as.character(round(as.numeric(mass),2))),
                 data=as.data.frame(plot_loadings) %>% rownames_to_column('mass'))

p2

ggsave("pca.png", plot=p1, device="png", dpi=100, path=results_folder,
       bg="white")
ggsave("pca.svg", plot=p1, device="svg", dpi=100, path=results_folder,
       bg="white")
ggsave("pca_loads.png", plot=p2, device="png", dpi=100, path=results_folder,
       bg="white")
ggsave("pca_loads.svg", plot=p2, device="svg", dpi=100, path=results_folder,
       bg="white")

```

## Separate species
```{r}
# Separate sheep and goat matrices
mat_sheep = brSVA[, uoc_metadata$species == 'Sheep']
mat_goat = brSVA[, uoc_metadata$species == 'Goat']
#Annotations for sheep and goat
uoc_metadata_sheep = uoc_metadata %>% filter(species == 'Sheep')
uoc_metadata_goat = uoc_metadata %>% filter(species == 'Goat')

annot_sheep = uoc_metadata_sheep %>%
  column_to_rownames('sample.name') %>%
  as.data.frame() %>%
  select(species)
cl_sheep = paste0(uoc_metadata_sheep$folia, uoc_metadata_sheep$side, '-', uoc_metadata_sheep$group, collapse = NULL)

annot_goat = uoc_metadata_goat %>%
  column_to_rownames('sample.name') %>%
  as.data.frame() %>%
  select(species)
cl_goat = paste0(uoc_metadata_goat$folia, uoc_metadata_goat$side, '-', uoc_metadata_goat$group, collapse = NULL)

```

### PCA by species

```{r}


sheep_pca = prcomp(t(mat_sheep))
sheep_pca_x = as.data.frame(sheep_pca$x)
sheep_pca_x = cbind(sheep_pca_x, uoc_metadata_sheep)
sheep_pca_x = sheep_pca_x %>% mutate(
  label = ifelse(folia==89|folia==1|folia==8|folia==97,
                 paste0(folia, side, '-', group, collapse = NULL),
                 ""))

perc_var = round(sheep_pca$sdev^2*100/sum(sheep_pca$sdev^2), 2)

# PC1 vs PC2
sheep_pca_plot = ggplot(sheep_pca_x) +
  geom_point(mapping = aes (x=PC1, y=PC2, color=species),
             size=5, shape=16, stroke=2) +
  xlab(paste0('PC1. ', perc_var[1], "% variance explained")) +
  ylab(paste0('PC2. ', perc_var[2], "% variance explained")) +
  scale_color_manual(values= c(Sheep = "#DBBD68")) + 
  geom_text(mapping = aes(x=PC1, y=PC2, label=as.character(group)),
            color="white", size=3) +
  geom_text_repel(mapping = aes(x=PC1, y=PC2, label=label),
            color="black", size=4) +
  theme_cowplot() +
  ggtitle("Sheep")
sheep_pca_plot




goat_pca = prcomp(t(mat_goat))
goat_pca_x = as.data.frame(goat_pca$x)
goat_pca_x = cbind(goat_pca_x, uoc_metadata_goat)
goat_pca_x = goat_pca_x %>% mutate(
  label = ifelse(folia==16|folia==9,
                 paste0(folia, side, '-', group, collapse = NULL),
                 ""))
perc_var = round(goat_pca$sdev^2*100/sum(goat_pca$sdev^2), 2)

# PC1 vs PC2
goat_pca_plot = ggplot(goat_pca_x) +
  geom_point(mapping = aes (x=PC1, y=PC2, color=species),
             size=5, shape=16, stroke=2) +
  xlab(paste0('PC1. ', perc_var[1], "% variance explained")) +
  ylab(paste0('PC2. ', perc_var[2], "% variance explained")) +
  scale_color_manual(values= c(Goat = "#CC6D68")) + 
  geom_text(mapping = aes(x=PC1, y=PC2, label=as.character(group)),
            color="white", size=4) +
  geom_text_repel(mapping = aes(x=PC1, y=PC2, label=label),
            color="black", size=4) +
  theme_cowplot() +
  ggtitle("Goat")
goat_pca_plot



ggsave("pca_sheep.png", plot=sheep_pca_plot, device="png", dpi=100, path=results_folder,
       bg="white")
ggsave("pca_sheep.svg", plot=sheep_pca_plot, device="svg", dpi=100, path=results_folder,
       bg="white")
ggsave("pca_goat.png", plot=goat_pca_plot, device="png", dpi=100, path=results_folder,
       bg="white")
ggsave("pca_goat.svg", plot=goat_pca_plot, device="svg", dpi=100, path=results_folder,
       bg="white")

```

### Heatmap by species

```{r, fig.width = 9, fig.height = 7}
#Heatmap for sheep

heatmap_sheep = pheatmap(
  t(mat_sheep),color=col2(8), treeheight_col = 0, show_colnames = F,
  clustering_method = 'ward.D2', labels_row = cl_sheep, cutree_row=3,
  fontsize = 8, main="Sheep")
heatmap_sheep
ggsave("heatmap_sheep.png", plot=heatmap_sheep, device="png", dpi=100,
       path=results_folder, width = 9, height = 7)
ggsave("heatmap_sheep.svg", plot=heatmap_sheep, device="svg", dpi=100,
       path=results_folder, width = 9, height = 7)

#Heatmap for goat

heatmap_goat = pheatmap(
  t(mat_goat), color=col2(8), treeheight_col = 0, show_colnames = F,
  clustering_method = 'ward.D2', labels_row = cl_goat, cutree_row=2,
  main="Goat")
heatmap_goat
ggsave("heatmap_goat.png", plot=heatmap_goat, device="png", dpi=100,
       path=results_folder, width = 9, height = 7)
ggsave("heatmap_goat.svg", plot=heatmap_goat, device="svg", dpi=100,
       path=results_folder, width = 9, height = 7)
```


## Statistical testing on each peptide
Here we build an statistical model for our data that includes the contribution of the species and the book each sample beolongs to.

For sample $i$, and feature peak $j$, the observed intensity signal $y$ is:

$y_{ij} = \alpha_j x_{ij} + \beta_j x_{ij} + \gamma_j x_{ij}$, 

where $\alpha_j$ is the species contribution of the peak $j$ and $\beta_j$ is for the book. The model also accounts for the batch, through parameter $\gamma$.

In the process we follow the following steps:

1. Fit model to the data to estimate all the $\alpha_i$ and $\beta_i$ parameters.
2. Define the contrasts we are interested in, in this case the difference between _sheep_ and _goat_ and the two books.
3. Fit the contrasts: perform the statistical testing to the defined contrasts and calcualte the size and and significance of the contrasts. Adjust p-values for multi-testing.

The results are sumarised in a volcano plot, showing the size and significance of the diferences.
Finally we do several diagnostic plots to determine if the assumptions are fulfilled.

```{r}
# Models for peptides testing
LB_Bmod = model.matrix(~0+species+book+batch, data=uoc_metadata)

# fit model on batch corrected data
fit.LBB = lmFit(brSVA, LB_Bmod)

# Create contrasts
LBB.contrast = makeContrasts(
  species_contrast = 'speciesSheep - speciesGoat',
  book_contrast = 'book2',
  levels = LB_Bmod
)
```

### Test and plot

```{r}
test.contrasts = function(fit, contr){
  fit2 = contrasts.fit(fit, contr)
  fit2 = eBayes(fit2, robust=T, trend=T)
  return(fit2)
}

diag_plots = function(fit2, base_main){
  par(mfrow=c(3,1))
  # SA Plot: avg vs dispersion
  plotSA(fit2, col=c('black', 'red'), cex=1.5, main='SA plot')
  # MA Plot: avg vs log2FC
  plotMA(fit2, cex=1.5, main='MA plot', status=p.adjust(fit2$F.p.value, method='BH')<0.05,
         value=T, legend=F)
  # p.values distriburtion
  hist(fit2$F.p.value, breaks=20, xlab='p-value',
       main='p-value distribution')
  title(base_main, line=1, outer = T, cex=2)
}

volc_plot = function(fit2, contr, base_main){
  # volcano plot per contrast coef: log2FC vs -log10pval
  # Also return a table with the data
  volcpl = list()
  ctrs = list()
  for (coef_idx in 1:length(colnames(contr))){
    volcdf = data.frame(mass=names(fit2$coefficients[,coef_idx]),
                        logFC = fit2$coeff[,coef_idx],
                        adj.p.val = p.adjust(fit2$p.value[,coef_idx], 'BH'),
                        p.val = fit2$p.value[,coef_idx],
                        stringsAsFactors = F)
    volcdf = volcdf %>%  mutate(l = ifelse(volcdf$adj.p.val <= 0.05, volcdf$mass, ''))
    DEstatus = cut(volcdf$adj.p.val, c(0,0.05,1))
    levels(DEstatus) = c('(0,0.05]','(0.05,1]')
    volcdf$DEstatus = DEstatus
    
    v = ggplot(volcdf, mapping = aes(x=logFC, y=-log10(adj.p.val))) +
      geom_point(mapping = aes(color=DEstatus)) +
      theme_cowplot() +
      ylab("-log10(FDR)") +
      geom_hline(yintercept = -log10(0.05), color='red') +
      scale_color_manual(name='FDR',
                         values=c('(0,0.05]'='#00BFC4','(0.05,1]'='#F8766D'),
                         na.value='grey66',
                         labels=c('FDR<=0.05','FDR>0.05')) +
      geom_text_repel(aes(label=l), color='black', size=3, max.overlaps = 20) +
      ggtitle(colnames(fit2$coeff)[coef_idx])
    
    v = v + theme(legend.position = 'right')

    ctrs[[colnames(fit2$coeff)[coef_idx]]] = volcdf
    volcpl[[coef_idx]] = v 
  }
  pl = ggarrange(plotlist = volcpl, nrow=1, ncol=length(colnames(contr)),
                common.legend = TRUE)
#   table = topTable(fit2, p.value=0.05, number=length(fit2$p.value))
#   table %>% rownames_to_column('mass') %>% write_csv(paste(base_file, '_table.csv', sep=''))
  a = list(ctrs, pl)
  return(a)
}
```

```{r, fig.width = 6, fig.height = 7}
# Test contrasts and plot
fit2.LBB = test.contrasts(fit.LBB, LBB.contrast)
diag_plots(fit2.LBB, 'SV corrected data, batch as factor.')
png(file.path(results_folder, 'model_diag_plots.png'),
    width=8, height=9, units='in', res=100)
diag_plots(fit2.LBB, 'SV corrected data, batch as factor.')
dev.off()
```

```{r, fig.width = 11, fig.height = 7}
a = volc_plot(fit2.LBB, LBB.contrast,'SV corrected data, batch as factor.')
tables.LBB = a[[1]]
vcpl = a[[2]]
vcpl
ggsave('volcano_plot.png', vcpl, path = results_folder, dpi=100, bg="white")
species.table = tables.LBB$species_contrast
book.table = tables.LBB$book_contrast

species_peps = species.table %>%
  filter(adj.p.val<=0.05) %>%
  arrange(adj.p.val) %>%
  dplyr::pull(mass) %>%
  as.numeric()

species.table %>%
  filter(adj.p.val<=0.05) %>%
  arrange(adj.p.val) %>%
  dplyr::select(-DEstatus)

book_peps = book.table %>%
  filter(adj.p.val<=0.05) %>%
  arrange(adj.p.val) %>%
  dplyr::pull(mass) %>%
  as.numeric()
book.table %>%
  filter(adj.p.val<=0.05) %>%
  arrange(adj.p.val) %>%
  dplyr::select(-DEstatus)

```

## Discriminant analysis

Then we use discriminant analysis using binary predictors, implemented in the package _binda_ again to detect peptides contributing to differences in the spectra that can be associated to the species or the book.


```{r, fig.width = 5, fig.height = 5}
colnames(featMatrix) <-
  round(as.double(colnames(featMatrix)),2)
top_rank = 20

# For species
thresh = optimizeThreshold(X=featMatrix, L=uoc_metadata$species, verbose=FALSE)
binMat = dichotomize(X=featMatrix, thresh)
bindar = binda.ranking(Xtrain=binMat, L=uoc_metadata$species, verbose=FALSE)
png(file.path(results_folder, 'bindar_species.png'),
    width=8, height=9, units='in', res=72)
plot(bindar, top=top_rank, arrow.col="black", ylab="Peaks (m/z)")
dev.off()
plot(bindar, top=top_rank, arrow.col="black", ylab="Peaks (m/z)")
attr(bindar, 'class') = 'matrix'
bindar_species = as_tibble(bindar, rownames='mass')
```

```{r, fig.width = 5, fig.height = 6}
# For goat
thresh = optimizeThreshold(X=t(mat_goat), L=uoc_metadata_goat$cluster, verbose=FALSE)
binMat = dichotomize(X=t(mat_goat), thresh)
bindar = binda.ranking(Xtrain=binMat, L=uoc_metadata_goat$cluster, verbose=FALSE)
png(file.path(results_folder, 'bindar_goat.png'),
    width=8, height=9, units='in', res=72)
plot(bindar, top=top_rank, arrow.col="black", ylab="Peaks (m/z)",
     main="Top 20 ranking masses for goat clusters")
dev.off()
plot(bindar, top=top_rank, arrow.col="black", ylab="Peaks (m/z)",
     main="Top 20 ranking masses for goat clusters")
attr(bindar, 'class') = 'matrix'
bindar_goat = as_tibble(bindar, rownames='mass') %>%
  mutate(mass = as.numeric(mass), score=as.numeric(score),
         t.1=as.numeric(t.1), t.2=as.numeric(t.2)) %>%
  select(-idx)
```

We need to remove the single sample from the sheep subset. It is sample UoC24.

```{r, fig.width = 5, fig.height = 6}
# For sheep
# Remove sample from cluster 3
uoc_metadata_sheep_2 = uoc_metadata_sheep %>% filter(!sample.name=="UoC24")
mat_sheep_2 = mat_sheep[, colnames(mat_sheep) != "UoC24"]

thresh = optimizeThreshold(X=t(mat_sheep_2), L=uoc_metadata_sheep_2$cluster, verbose=FALSE)
binMat = dichotomize(X=t(mat_sheep_2), thresh)
png(file.path(results_folder, 'bindar_sheep.png'),
    width=8, height=9, units='in', res=72)
bindar = binda.ranking(Xtrain=binMat, L=uoc_metadata_sheep_2$cluster, verbose=FALSE)
plot(bindar, top=top_rank, arrow.col="black", ylab="Peaks (m/z)",
     main="Top 20 ranking masses for sheep clusters")
dev.off()
plot(bindar, top=top_rank, arrow.col="black", ylab="Peaks (m/z)",
     main="Top 20 ranking masses for sheep clusters")
attr(bindar, 'class') = 'matrix'
bindar_sheep = as_tibble(bindar, rownames='mass') %>%
  mutate(mass=as.numeric(mass), score=as.numeric(score),
         t.4=as.numeric(t.4), t.5=as.numeric(t.5)) %>%
  select(-idx)
```

## Find non-collagenous peptides masses

```{r}
library(bacollite)
```


```{r}
parse_seqs = function(species, chain, id, sequence){
  pseq = parse.seq(sequence, max.missed.cleaves = 1)
  pseq$chain=chain
  pseq$species=species
  pseq
}
# Read sequences
collagen = read_csv(file.path(data_folder, 'collagen_seqs.txt'))
# Digest sequences
peptides = do.call(mcmapply, c(list('FUN'=parse_seqs, 'mc.cores'=4L,
                                    'SIMPLIFY'=F, 'USE.NAMES'=F),
                               as.list(collagen)))
peptides = do.call(bind_rows, peptides) %>% as_tibble()

# peptides = peptides[peptides$nglut==0,]
peptides = peptides[!duplicated(select(peptides, -species, -seqpos)),] %>%
  select(-species)

top_sheep = bindar_sheep %>% mutate(cluster=ifelse(t.4>0, 4, 5))
top_sheep %>% write_csv(file.path(results_folder, "binda_all_sheep.csv"))
top_sheep = top_sheep[1:top_rank,] %>% arrange(mass) %>%
  write_csv(file.path(results_folder, paste0("binda_top", top_rank,"_sheep.csv")))
top_sheep = top_sheep %>% select(-t.4, -t.5)

top_goat = bindar_goat %>% mutate(cluster=ifelse(t.1>0, 1, 2))
top_goat %>% write_csv(file.path(results_folder, "binda_all_goat.csv"))
top_goat = top_goat[1:top_rank,] %>% arrange(mass) %>%
  write_csv(file.path(results_folder, paste0("binda_top", top_rank,"_goat.csv")))
top_goat = top_goat %>% select(-t.1, -t.2)

```


### First round search
```{r}

match_binda = function(top_mass, peptides_masses, tol){
  match.closest(top_mass, peptides_masses, tol)
}
tol = 0.6

goat_idx_match = do.call(c,
  lapply(top_goat$mass,
       match_binda,
       sort(peptides$mass1),
       tol)
)

top_goat[is.na(goat_idx_match),]


sheep_idx_match = do.call(c,
  lapply(top_sheep$mass,
       match_binda,
       sort(peptides$mass1),
       tol)
)

top_sheep[is.na(sheep_idx_match),]

```

## Second round search


First we get the matching theoretical peptides will be aligned.
We are also including deamidated od undeamidated versions of the peptides to
extend the search. Sometimes the other version alignes better to the spectra.
```{r}

match_seq = function(p, top_binda, tol) {
  idx = do.call(
    c,
    lapply(
      p$mass1,
      function(x, tol){
        match.closest(x, top_binda$mass, tol)
      },
      tol
    )
  )
  if (all(is.na(idx))){
    return(NULL)
  } else {
    return(cbind(p, top_binda[idx,]))
  }
}

tol = 0.6

peptides_split = group_split(peptides, seq, nhyd)

match_sheep = mclapply(peptides_split,
         match_seq,
         top_sheep, tol) %>%
  bind_rows() %>% as_tibble() %>%
  mutate(species="Sheep")


match_goat = mclapply(peptides_split,
         match_seq,
         top_goat, tol) %>%
  bind_rows() %>% as_tibble() %>%
  mutate(species="Goat")

```

Plot mass-errors
```{r}

comb_match = bind_rows(match_sheep, match_goat) %>% as_tibble() %>%
  filter(!is.na(mass)) %>% mutate(mass_error = mass1-mass)


ggplot(comb_match) +
  geom_point(aes(x=mass, y=mass_error, color=species))
```

Use bacollite to read-in the spectra
```{r}
# Read spectra

read_bacollite = function(sample, folder, fext, repsep="_"){
  froot = paste0(folder, '/', sample)
  spots = paste0(repsep, c(1,2,3))
  s = load.sample(froot, sample, spots, fext)
  s$name = sample
  return(s)
}


bl_unproc = lapply(
  uoc_metadata$sample.name,
  read_bacollite,
  file.path(data_folder, 'UoC_txt'),
  fext='.txt')

names(bl_unproc) = uoc_metadata$sample.name

# Divide species
bl_goat = bl_unproc[uoc_metadata %>% filter(species=="Goat") %>% pull(sample.name)]

bl_sheep = bl_unproc[uoc_metadata %>% filter(species=="Sheep") %>% pull(sample.name)]



```

Functions to align theoretical peptides to samples
```{r}
ps_fit = function(x, m, f){
  a = ms_fit(peptides = m, sample = x, laglim=0.6,
                       doplot=F, force=T, gauss = 0.2)
  a = as_tibble(a)
  a$sample_name = x$name
  a
}

corr_pept = function(seq, nhyd, nglut, mass1, seqpos, missed.cleaves, chain,
                     mass, score, cluster, species,
                     bl, laglim, f) {
  m = data.frame(
        seq=seq, nhyd=nhyd, nglut=nglut, mass1=mass1, seqpos=seqpos,
        missed.cleaves=missed.cleaves,  chain=chain, mass=mass,
        score=score, cluster=cluster, species=species
      )
  hit_bac = mclapply(
    bl, ps_fit, m, f, mc.cores = 4
  )
  hit_bac = bind_rows(hit_bac)
  
  cl_bac = corlim_data(hit_bac, fn_laglim = laglim)
  bind_cols(m, cl_bac)
}
```

Align peptides from goat clusters to goat samples
```{r}
bac_cor_goat = do.call(
    mapply,
    c(list('FUN'=corr_pept, 'SIMPLIFY'=F, 'USE.NAMES'=F,
           'MoreArgs'=list('bl'=bl_goat, 'laglim'=0.3,
                           'f'=file.path(results_folder, 'bacollite_figs'))),
      as.list(match_goat)))

bac_cor_goat = bind_rows(bac_cor_goat)

bac_cor_goat = bac_cor_goat %>% as_tibble() %>%
  mutate(nglut=as.factor(nglut),
         nhyd=as.factor(nhyd),
         cluster=as.factor(cluster),
         mass = as.factor(mass),
         mass1 = as.factor(round(mass1, 2)))

```

Plot correlation results
```{r, fig.width=7, fig.height=12}
threshold_plot = list(
  ggplot(bac_cor_goat %>% filter(cluster==1)) +
    geom_line(aes(x=cl, y=nh, color=nhyd,
                  group=interaction(nhyd, nglut, sep=' - '))) +
    geom_point(aes(x=cl, y=nh, color=nhyd, shape=nglut,
                   group=interaction(nhyd, nglut, sep=' - '))) +
    facet_wrap(vars(mass), ncol=3) +
    theme(axis.title.x = element_blank()) +
    ylab("# hits") +
    ggtitle('Goat cluster 1'),
  
  ggplot(bac_cor_goat %>% filter(cluster==2)) +
    geom_line(aes(x=cl, y=nh, color=nhyd,
                  group=interaction(nhyd, nglut, sep=' - '))) +
    geom_point(aes(x=cl, y=nh, color=nhyd, shape=nglut,
                   group=interaction(nhyd, nglut, sep=' - '))) +
    facet_wrap(vars(mass), ncol=3) +
    theme(axis.title.x = element_blank()) +
    ylab("# hits") +
    ggtitle('Goat cluster 2'),
  
  ggplot(bac_cor_goat %>% filter(is.na(cluster))) +
    geom_line(aes(x=cl, y=nh, color=nhyd,
                  group=interaction(nhyd, nglut, sep=' - '))) +
    geom_point(aes(x=cl, y=nh, color=nhyd, shape=nglut,
                   group=interaction(nhyd, nglut, sep=' - '))) +
    facet_wrap(vars(mass1), ncol=3) +
    xlab('Cross-correlation threshold') +
    ylab("# hits") +
    ggtitle('Goat extended')
)

pl = ggarrange(plotlist = threshold_plot, nrow = 3,
               heights = c(1, 1.5, 2))
pl

ggsave('bcl_goat.png', pl, path = results_folder, height = 12, width = 7)
```

```{r}
bac_cor_sheep = do.call(
    mapply,
    c(list('FUN'=corr_pept, 'SIMPLIFY'=F, 'USE.NAMES'=F,
           'MoreArgs'=list('bl'=bl_sheep, 'laglim'=0.3,
                           'f'=file.path(results_folder, 'bacollite_figs'))),
      as.list(match_sheep)))

bac_cor_sheep = bind_rows(bac_cor_sheep)

bac_cor_sheep = bac_cor_sheep %>% as_tibble() %>%
  mutate(nglut=as.factor(nglut),
         nhyd=as.factor(nhyd),
         cluster=as.factor(cluster),
         mass = as.factor(mass),
         mass1 = as.factor(round(mass1, 2)))
```


```{r, fig.width=7, fig.height=12}

threshold_plot = list( 
  ggplot(bac_cor_sheep %>% filter(cluster==4)) +
    geom_line(aes(x=cl, y=nh, color=nhyd,
                  group=interaction(nhyd, nglut, sep=' - '))) +
    geom_point(aes(x=cl, y=nh, color=nhyd, shape=nglut,
                   group=interaction(nhyd, nglut, sep=' - '))) +
    facet_wrap(vars(mass), ncol=4) +
    theme(axis.title.x = element_blank()) +
    ylab("# hits") +
    ggtitle('Sheep cluster 4'),
  
  ggplot(bac_cor_sheep %>% filter(cluster==5)) +
    geom_line(aes(x=cl, y=nh, color=nhyd,
                  group=interaction(nhyd, nglut, sep=' - '))) +
    geom_point(aes(x=cl, y=nh, color=nhyd, shape=nglut,
                   group=interaction(nhyd, nglut, sep=' - '))) +
    facet_wrap(vars(mass), ncol=4) +
    theme(axis.title.x = element_blank()) +
    ylab("# hits") +
    ggtitle('Sheep cluster 5'),
  
  ggplot(bac_cor_sheep %>% filter(is.na(cluster))) +
    geom_line(aes(x=cl, y=nh, color=nhyd,
                  group=interaction(nhyd, nglut, sep=' - '))) +
    geom_point(aes(x=cl, y=nh, color=nhyd, shape=nglut,
                   group=interaction(nhyd, nglut, sep=' - '))) +
    facet_wrap(vars(mass1), ncol=4) +
    xlab('Cross-correlation threshold') +
    ylab("# hits") +
    ggtitle('Sheep extended')
)
pl = ggarrange(plotlist = threshold_plot, nrow = 3,
               heights = c(1, 1, 2))
pl

ggsave('bcl_sheep.png', pl, path = results_folder, height = 12, width = 7)
```

